Designing Types for R, Empirically

### 2020-05-16 paper/main.tex
Updated 4.3

In terms of usability, our checker is enabled automatically whenever a
package is loaded in R. So, a simple invocation of \code{library(contractr)}
causes contracts to be injected in all packages.  On loading, \contractr
scans all packages in the user's workspace and inserts contracts in
functions for which type signatures are available. Then \contractr sets up
package load hooks which are executed when new packages are loaded.
Furthermore, \contractr automatically removes contracts from all functions
and restores them to their original state when it is unloaded.
The type signatures can be provided externally, in a file, thus avoiding the
need to change the source code of checked packages.  Type declarations can
also be written alongside a \code{@type} section inside \roxygen function
comment blocks. \roxygen is an R package that enables authors to add
documentation. The documentation sections are processed by \roxygen before
package build and a registered method for each custom tag section is
invoked with the section data and documented object. \contractr uses this
mechanism to register a hook for a custom \code{@type} tag. 
The \contractr API allows users to explicitly insert contracts during
interactive development by supplying the type signature as a string.
Contracts can be removed by calling \code{remove\_contract}.  Contracts can
be selectively enabled or disabled for code blocks by wrapping them in calls
to \code{ignore\_contracts} and \code{capture\_contracts}.  These functions
return a data frame that contains all the information about failed and
successful contract assertions in the wrapped code blocks.  Multiple
failures can introduce noise; this is avoided by setting
\code{contractr.severity} to \code{'silence'}.  \contractr still performs
type-checking so that results can be obtained by calling
\code{get\_contracts}. Setting \code{contractr.severity} flag to
\code{'error'} turns the type-checking warnings to errors and
halt the program.

>>>>

In terms of usability, our checker is enabled automatically whenever a package
is loaded in R. So, a simple invocation of \code{library(contractr)} causes
contracts to be injected in all packages. On loading, \contractr scans all
packages in the user's workspace and inserts contracts in functions for which
type signatures are available. Then \contractr sets up package load hooks which
are executed when new packages are loaded. Furthermore, \contractr automatically
removes contracts from all functions and restores them to their original state
when it is unloaded. The type signatures can be provided externally, in a file,
thus avoiding the need to change the source code of checked packages. Type
declarations can also be written directly in the adjunct function comments using
\roxygen annotation tags. \contractr introduces a new use \code{@type} tag. When
a package these type signatures are extracted and stored in the appropriate
location. The following is an example of a type annotation using \roxygen:

\begin{lstlisting}
#' @type <chr> => int
#' @export
file_size <- function(f) { ... }
\end{lstlisting}

The \contractr provides an extensive API that allows users to explicitly insert
contracts during interactive development by supplying the type signature as a
string, selectively disable contract checking and inspecting type assertions.

% Contracts can be removed by calling \code{remove\_contract}.  Contracts can
% be selectively enabled or disabled for code blocks by wrapping them in calls
% to \code{ignore\_contracts} and \code{capture\_contracts}.  These functions
% return a data frame that contains all the information about failed and
% successful contract assertions in the wrapped code blocks.  Multiple
% failures can introduce noise; this is avoided by setting
% \code{contractr.severity} to \code{'silence'}.  \contractr still performs
% type-checking so that results can be obtained by calling
% \code{get\_contracts}. Setting \code{contractr.severity} flag to
% \code{'error'} turns the type-checking warnings to errors and
% halt the program.
